\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3.2\baselineskip}

### 2017-03-24 paper/ms.tex
small abstract edits

The growing field of large-scale time domain astronomy requires methods for
probabilistic inference that are computationally tractable, even with large
datasets.
We present a method for Gaussian Process regression in one dimension with a
computational cost that scales linearly with the size of the dataset.
This method exploits structure in the problem when the covariance function is
expressed as a mixture of complex exponentials, without requiring evenly
spaced observations or uniform noise.
This form of covariance arises naturally when the process is a mixture of
stochastically-driven damped harmonic oscillators~--~providing a physical
motivation for and interpretation of this choice~--~but we also demonstrate
that it is effective in many other cases.
We present a mathematical description of the method, the details of the
implementation, and a comparison to existing scalable Gaussian Process methods.
We demonstrate the method by applying it to simulated and real astronomical
time series datasets.
These demonstrations are examples of probabilistic inference of stellar
rotation periods, asteroseismic oscillation spectra, and transiting planet
parameters.
In all of these cases, we infer posterior distributions over models for the
power spectrum of the process that generated these irregularly sampled time
series without computing a standard algebraic estimator.
The method is flexible, fast, and most importantly, interpretable, with a wide
range of potential applications within astronomical data analysis and beyond.
We provide well-tested and documented open-source implementations of this
method in \project{C++}, \project{Python}, and \project{Julia}.

>>>>

The growing field of large-scale time domain astronomy requires methods for
probabilistic inference that are computationally tractable, even with large
datasets.
Gaussian Processes are a popular class of models used for this purpose but,
since the computational cost scales as the cube of the number of data points,
their application has been limited to relatively small datasets.
In this paper, we present a method for Gaussian Process regression in
one-dimension where the computational requirements scale linearly with the
size of the dataset.
This method exploits structure in the problem when the covariance function is
expressed as a mixture of complex exponentials, without requiring evenly
spaced observations or uniform noise.
This form of covariance arises naturally when the process is a mixture of
stochastically-driven damped harmonic oscillators~--~providing a physical
motivation for and interpretation of this choice~--~but we also demonstrate
that it is effective in many other cases.
We present a mathematical description of the method, the details of the
implementation, and a comparison to existing scalable Gaussian Process methods.
We demonstrate the method by applying it to simulated and real astronomical
time series datasets.
These demonstrations are examples of probabilistic inference of stellar
rotation periods, asteroseismic oscillation spectra, and transiting planet
parameters.
In all of these cases, we infer posterior distributions over models for the
power spectrum of the process that generated these irregularly sampled time
series without computing a standard algebraic estimator.
The method is flexible, fast, and most importantly, interpretable, with a wide
range of potential applications within astronomical data analysis and beyond.
We provide well-tested and documented open-source implementations of this
method in \project{C++}, \project{Python}, and \project{Julia}.
